---
title: "Analysis of COVID-19 Apple Mobility Data"
author: "Xienyam Chiu"
date: "10/3/2021"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## Setup

```{r set-parameters}
# set the state to work with
state_to_analyze <- "South Carolina"

# sanitize the state name for output files
state_no_spaces <- gsub("\\s", "_", state_to_analyze)

# set the name of the global date file to use
file_to_process <- "data/data_raw/applemobilitytrends-2021-09-20.csv"

# extract core name for output file naming
core_name <- tools::file_path_sans_ext(basename(file_to_process))
```

```{r load-packages-functions}
# load the packages needed
library("dplyr")
library("readr")
library("ggplot2")
library("tidyr")
library("lubridate")
library("devtools")

# load the functions needed
source("code/functions/state_subset_mobility_data.R")
source("code/functions/state_transportation_tally.R")
source("code/functions/state_data_to_long.R")
```

## Background and Objectives 

This rmarkdown file will use `ggplot2` [@ggplot2] and `dplyr` [@dplyr] packages to analyze the Apple maps mobility data from the time before the COVID-19 pandemic through the initial shutdowns and the continuing changes over the course of the pandemic.

The goal will be to create a subset to a particular state, tally up the number of cities and counties in that state with mobililty data, and then to make several plots to visual the data subsets.

The state that this document will analyze is `r state_to_analyze`.

## Data Processing 

```{r subset-state}
# subset out to the chose state using the variables set above
state_subset <- state_subset_mobility_data(
  input_file_name = file_to_process,
  state_to_subset = state_to_analyze)
```

```{r make-subset-long}
# convert subsetted state date from wide to long format
long_subset <- state_data_to_long(
  file = paste0("output/subsetted_states_wide/",
                core_name,
                "_",
                state_no_spaces,
                ".csv"))
```

```{r tallied_subset}
# tally up the number of cities and counties in the state that have mobility
# data available
tallied_subset <- state_transportation_tally(
  path_name = core_name,
  state = state_to_analyze)
```

## Figures

```{r plot-tallied}
# create a grouped barplot of the tallied data to inclue in the knitted
# document and save that figure to the output/figures directory as a png file
tallied_plot <- ggplot(data = tallied_subset,
                       aes(x = geo_type,
                           y = n,
                           fill = transportation_type)) +
  geom_col(position = position_dodge()) +
  labs(title = paste("Number of cities and/or counties in",
                     state_to_analyze,
                     "with mobility data"),
       x = "Type of area",
       y = "Number of areas")

ggsave(plot = tallied_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_tallied_plot.png"))
tallied_plot
```

```{r plot-time-series}
# produce a line plot of relative driving mobility across the state
# with date on the X axis and mean relative mobility on the Y axis
timeseries_plot <- long_subset %>%
  filter(transportation_type == "driving") %>%
  group_by(date) %>%
  summarize(mean_mobility = mean(relative_mobility)) %>%
  ggplot(aes(x = lubridate::ymd(date),
             y = mean_mobility)) +
  geom_line() +
  labs(title = paste("Statewide mean relative mobility driving levels in",
                     state_to_analyze,
                     "during COVID"),
       x = "Date",
       y = "Mean relative mobility")
ggsave(plot = timeseries_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_timeseries_plot.png"))

timeseries_plot
```

## Session Info

```{r session-info}
devtools::session_info()
```

## Sources Cited

